---
title: "Utility"
author: "Cameron Flores"
date: "10/24/2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(qwraps2)
library(knitr)
library(ggplot2)
library(productplots)
library(brms)
library(bayesplot)
## read data set
survey <- data.frame(readr::read_csv(file = "survey.csv"))

survey <- survey %>%
select("age", "gender", "race", "employ", "educ", "alc", "drug", "cann", "cash", "insurance", "region")

survey$age[survey$age==1] <- 2
survey$age[survey$age==0] <- 1

survey$gender[survey$gender==1] <- 2
survey$gender[survey$gender==0] <- 1

survey$alc[survey$alc==1] <- 2
survey$alc[survey$alc==0] <- 1

survey$drug[survey$drug==1] <- 2
survey$drug[survey$drug==0] <- 1

survey$cann[survey$cann==1] <- 2
survey$cann[survey$cann==0] <- 1

survey$cash[survey$cash==1] <- 2
survey$cash[survey$cash==0] <- 1

survey$insurance[survey$insurance==1] <- 2
survey$insurance[survey$insurance==0] <- 1

survey$age <- as.factor(survey$age)
survey$gender <- as.factor(survey$gender)
survey$alc <- as.factor(survey$alc)
survey$drug <- as.factor(survey$drug)
survey$cann <- as.factor(survey$cann)
survey$cash <- as.factor(survey$cash)

survey
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
library(NPBayesImputeCat)

dj <- c(2, 2, 4, 5, 6, 2, 2, 2, 2, 2, 5)
r <- ncol(survey)
# make sure all variables are categorical
for (j in 1:r){
  survey[, j] <- factor(survey[, j], levels = 1:dj[j])
}

#select random sample of three rows from data frame
survey <- survey[sample(1:nrow(survey), 5000), ]
```

```{r size = "footnotesize", results = 'hide', message = FALSE, warning = FALSE}
model <- NPBayesImputeCat::CreateModel(X = survey,
                                       MCZ = NULL,
                                       K = 70,
                                       Nmax = 0,
                                       aalpha = 0.25,
                                       balpha = 0.25,
                                       seed = 221)
```

```{r message=FALSE, warning=FALSE, results='hide', size="footnotesize"}
m <- 5
survey_syn <- NPBayesImputeCat::DPMPM_nozeros_syn(X = survey,
                                                   dj = dj,
                                                   nrun = 5000,
                                                   burn = 500,
                                                   thin = 10,
                                                   K = 70,
                                                   aalpha = 0.25,
                                                   balpha = 0.25,
                                                   m = m,
                                                   vars = c("region", "alc", "drug", "cann", "cash"),
                                                   seed = 221,
                                                   silent = TRUE)
```

```{r size = "footnotesize", eval = FALSE, message = FALSE, warning = FALSE, echo=FALSE}
library(bayesplot)
kstar_MCMCdiag(kstar = survey_syn$kstar,
               nrun = 5000,
               burn = 500,
               thin = 10)
```

```{r eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 4, fig.width = 4, echo=FALSE}
NPBayesImputeCat::marginal_compare_all_syn(obsdata = survey,
                                           syndata = survey_syn$syndata,
                                           vars = "region")
```

```{r eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 4, fig.width = 4, echo=FALSE}
NPBayesImputeCat::marginal_compare_all_syn(obsdata = survey,
                                           syndata = survey_syn$syndata,
                                           vars = "alc")
```

```{r eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 4, fig.width = 4, echo=FALSE}
NPBayesImputeCat::marginal_compare_all_syn(obsdata = survey,
                                           syndata = survey_syn$syndata,
                                           vars = "drug")
```

```{r eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 4, fig.width = 4, echo=FALSE}
NPBayesImputeCat::marginal_compare_all_syn(obsdata = survey,
                                           syndata = survey_syn$syndata,
                                           vars = "cann")
```

```{r eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 4, fig.width = 4, echo=FALSE}
NPBayesImputeCat::marginal_compare_all_syn(obsdata = survey,
                                           syndata = survey_syn$syndata,
                                           vars = "cash")
```

```{r message=FALSE, warning=FALSE, results='hide', size="footnotesize"}
m <- 1
utility_survey_syn <- NPBayesImputeCat::DPMPM_nozeros_syn(X = survey,
                                                   dj = dj,
                                                   nrun = 5000,
                                                   burn = 500,
                                                   thin = 10,
                                                   K = 70,
                                                   aalpha = 0.25,
                                                   balpha = 0.25,
                                                   m = m,
                                                   vars = c("region", "alc", "drug", "cann", "cash"),
                                                   seed = 221,
                                                   silent = TRUE)
```

## Global Utility Evaluation (pMSE)
```{r}
n <- nrow(survey)
merged_data <- rbind(survey, 
                     survey_syn$syndata[[1]])
merged_data$S <- c(rep(0, n), rep(1, n))
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
log_reg <- stats::glm(formula = S ~ (as.factor(age) + as.factor(gender) + as.factor(employ) +
                        as.factor(race) + as.factor(educ) + as.factor(alc) + as.factor(drug) +
                        as.factor(cann) + as.factor(cash) + as.factor(insurance) +
                          as.factor(region))^2,
                      family = "binomial", 
                      data = merged_data)
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
pred <- stats::predict(log_reg,
                       newdata = merged_data)
probs <- exp(pred) / (1 + exp(pred))
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
pMSE <- 1 / (2 * n) * sum((probs - 1 / 2)^2)
pMSE
```
The calculated pMSE is small and close to 0, therefore, the logistic regression model cannot distinguish between the confidential and synthetic dataset, indicating a high level of utility of our synthetic dataset.

Now we do this for the other 4 and calculate the overall average pMSE
```{r size = "footnotesize", warning = FALSE, message = FALSE}
merged_data2 <- rbind(survey, 
                     survey_syn$syndata[[2]])
merged_data2$S <- c(rep(0, n), rep(1, n))
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
log_reg2 <- stats::glm(formula = S ~ (as.factor(age) + as.factor(gender) + as.factor(employ) +
                        as.factor(race) + as.factor(educ) + as.factor(alc) + as.factor(drug) +
                        as.factor(cann) + as.factor(cash) + as.factor(insurance) +
                          as.factor(region))^2,
                      family = "binomial", 
                      data = merged_data2)
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
pred2 <- stats::predict(log_reg2,
                       newdata = merged_data2)
probs2 <- exp(pred2) / (1 + exp(pred2))
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
pMSE2 <- 1 / (2 * n) * sum((probs2 - 1 / 2)^2)
pMSE2
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
merged_data3 <- rbind(survey, 
                     survey_syn$syndata[[3]])
merged_data3$S <- c(rep(0, n), rep(1, n))
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
log_reg3 <- stats::glm(formula = S ~ (as.factor(age) + as.factor(gender) + as.factor(employ) +
                        as.factor(race) + as.factor(educ) + as.factor(alc) + as.factor(drug) +
                        as.factor(cann) + as.factor(cash) + as.factor(insurance) +
                          as.factor(region))^2,
                      family = "binomial", 
                      data = merged_data3)
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
pred3 <- stats::predict(log_reg3,
                       newdata = merged_data3)
probs3 <- exp(pred3) / (1 + exp(pred3))
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
pMSE3 <- 1 / (2 * n) * sum((probs3 - 1 / 2)^2)
pMSE3
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
merged_data4 <- rbind(survey, 
                     survey_syn$syndata[[4]])
merged_data4$S <- c(rep(0, n), rep(1, n))
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
log_reg4 <- stats::glm(formula = S ~ (as.factor(age) + as.factor(gender) + as.factor(employ) +
                        as.factor(race) + as.factor(educ) + as.factor(alc) + as.factor(drug) +
                        as.factor(cann) + as.factor(cash) + as.factor(insurance) +
                          as.factor(region))^2,
                      family = "binomial", 
                      data = merged_data4)
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
pred4 <- stats::predict(log_reg4,
                       newdata = merged_data4)
probs4 <- exp(pred4) / (1 + exp(pred4))
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
pMSE4 <- 1 / (2 * n) * sum((probs4 - 1 / 2)^2)
pMSE4
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
merged_data5 <- rbind(survey, 
                     survey_syn$syndata[[5]])
merged_data5$S <- c(rep(0, n), rep(1, n))
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
log_reg5 <- stats::glm(formula = S ~ (as.factor(age) + as.factor(gender) + as.factor(employ) +
                        as.factor(race) + as.factor(educ) + as.factor(alc) + as.factor(drug) +
                        as.factor(cann) + as.factor(cash) + as.factor(insurance) +
                          as.factor(region))^2,
                      family = "binomial", 
                      data = merged_data5)
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
pred5 <- stats::predict(log_reg5,
                       newdata = merged_data5)
probs5 <- exp(pred5) / (1 + exp(pred5))
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
pMSE5 <- 1 / (2 * n) * sum((probs5 - 1 / 2)^2)
pMSE5
```

```{r}
(pMSE+pMSE2+pMSE3+pMSE4+pMSE5)/5
```

The average pMSE is small and close to 0 meaning the logistic regression cannot distinguish between the confidential and synthetic datasets, indicating a high level of utility of our synthetic datasets.
```{r}
merged_data
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
library(RSNNS)
library(caret)
```
```{r size = "footnotesize", warning = FALSE, message = FALSE}
#use merged_data
#shuffle the vector
data <- merged_data[sample(1:nrow(merged_data),length(1:nrow(merged_data))),1:ncol(merged_data)]

dummy <- dummyVars(" ~ .", data = data)
survey_mlp <- data.frame(predict(dummy, newdata = data ))

surveyValues <- survey_mlp[,1:11]

surveyTargets <- decodeClassLabels(survey_mlp[,12])

train_ind <- sample(seq_len(nrow(survey_mlp)), size = smp_size)

train <- survey_mlp[train_ind, ]

train_x <- train[,1:11]
train_y <- train[,12]

test <- survey_mlp[-train_ind, ]
test_x <- test[,1:11]
test_y <- test[,12]

# TRY CHANGING SPLIT FUNCTION
survey_mlp <- splitForTrainingAndTest(surveyValues, surveyTargets, ratio=0.15)
```


```{r size = "footnotesize", warning = FALSE, message = FALSE}
model <- mlp(train_x, 
             train_y, 
             size = c(10, 5), 
             initFunc = "Randomize_Weights",
             initFuncParams = c(-0.3, 0.3),
             learnFuncParams=c(0.1, 0),
             learnFunc = "Rprop",
             maxit = 100)

```


```{r size = "footnotesize", warning = FALSE, message = FALSE}

pred <- stats::predict(model, newdata = test_x)
probs <- exp(pred) / (1 + exp(pred))
pMSE <- 1 / (2 * n) * sum((probs - 1 / 2)^2)
pMSE

```
```{r size = "footnotesize", warning = FALSE, message = FALSE}
merged_data2 <- rbind(survey, 
                     survey_syn$syndata[[2]])
merged_data2$S <- c(rep(0, n), rep(1, n))

```
```{r size = "footnotesize", warning = FALSE, message = FALSE}
set.seed(1)
#use merged_data
#shuffle the vector
data2 <- merged_data2[sample(1:nrow(merged_data2),length(1:nrow(merged_data2))),1:ncol(merged_data2)]

dummy2 <- dummyVars(" ~ .", data = data2)
survey_mlp2 <- data.frame(predict(dummy2, newdata = data2 ))

surveyValues2 <- survey_mlp2[,1:11]

surveyTargets2 <- decodeClassLabels(survey_mlp2[,12])

train_ind2 <- sample(seq_len(nrow(survey_mlp2)), size = smp_size)

train2 <- survey_mlp2[train_ind2, ]

train_x2 <- train2[,1:11]
train_y2 <- train2[,12]

test2 <- survey_mlp2[-train_ind2, ]
test_x2 <- test2[,1:11]
test_y2 <- test2[,12]

survey_mlp2 <- splitForTrainingAndTest(surveyValues2, surveyTargets2, ratio=0.15)
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
set.seed(12)
model2 <- mlp(train_x2, 
             train_y2, 
             size = c(10, 5), 
             initFunc = "Randomize_Weights",
             initFuncParams = c(-0.3, 0.3),
             learnFuncParams=c(0.1, 0),
             learnFunc = "Rprop",
             maxit = 100)



```


```{r size = "footnotesize", warning = FALSE, message = FALSE}

pred2 <- stats::predict(model2, newdata = test_x2)
probs2 <- exp(pred2) / (1 + exp(pred2))
pMSE2 <- 1 / (2 * n) * sum((probs2 - 1 / 2)^2)
pMSE2

```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
#use merged_data
#shuffle the vector
set.seed(123)
data3 <- merged_data3[sample(1:nrow(merged_data3),length(1:nrow(merged_data3))),1:ncol(merged_data3)]

dummy3 <- dummyVars(" ~ .", data = data3)
survey_mlp3 <- data.frame(predict(dummy3, newdata = data3 ))

surveyValues3 <- survey_mlp3[,1:11]

surveyTargets3 <- decodeClassLabels(survey_mlp3[,12])

survey_mlp3 <- splitForTrainingAndTest(surveyValues3, surveyTargets3, ratio=0.15)
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
set.seed(2)
model3 <- mlp(survey_mlp3$inputsTrain, 
             survey_mlp3$targetsTrain, 
             size = c(10, 5), 
             initFunc = "Randomize_Weights",
             initFuncParams = c(-0.3, 0.3),
             learnFuncParams=c(0.1, 0),
             learnFunc = "Rprop",
             maxit = 100, 
             inputsTest=survey_mlp3$inputsTest, 
             targetsTest=survey_mlp3$targetsTest)



```


```{r size = "footnotesize", warning = FALSE, message = FALSE}
set.seed(23)
pred3 <- stats::predict(model3, newdata = survey_mlp3$inputsTest)
probs3 <- exp(pred3) / (1 + exp(pred3))
pMSE3 <- 1 / (2 * n) * sum((probs3 - 1 / 2)^2)
pMSE3

```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
#use merged_data
#shuffle the vector
set.seed(234)
data4 <- merged_data4[sample(1:nrow(merged_data4),length(1:nrow(merged_data4))),1:ncol(merged_data4)]

dummy4 <- dummyVars(" ~ .", data = data4)
survey_mlp4 <- data.frame(predict(dummy4, newdata = data4 ))

surveyValues4 <- survey_mlp4[,1:11]

surveyTargets4 <- decodeClassLabels(survey_mlp4[,12])

survey_mlp4 <- splitForTrainingAndTest(surveyValues4, surveyTargets4, ratio=0.15)
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
set.seed(3)
model4 <- mlp(survey_mlp4$inputsTrain, 
             survey_mlp4$targetsTrain, 
             size = c(10, 5), 
             initFunc = "Randomize_Weights",
             initFuncParams = c(-0.3, 0.3),
             learnFuncParams=c(0.1, 0),
             learnFunc = "Rprop",
             maxit = 100, 
             inputsTest=survey_mlp4$inputsTest, 
             targetsTest=survey_mlp4$targetsTest)


```


```{r size = "footnotesize", warning = FALSE, message = FALSE}
set.seed(34)
pred4 <- stats::predict(model4, newdata = survey_mlp4$inputsTest)
probs4 <- exp(pred4) / (1 + exp(pred4))
pMSE4 <- 1 / (2 * n) * sum((probs4 - 1 / 2)^2)
pMSE4

```


```{r size = "footnotesize", warning = FALSE, message = FALSE}
#use merged_data
#shuffle the vector
set.seed(345)
data5 <- merged_data5[sample(1:nrow(merged_data5),length(1:nrow(merged_data5))),1:ncol(merged_data5)]

dummy5 <- dummyVars(" ~ .", data = data5)
survey_mlp5 <- data.frame(predict(dummy5, newdata = data5 ))

surveyValues5 <- survey_mlp5[,1:11]

surveyTargets5 <- decodeClassLabels(survey_mlp5[,12])

survey_mlp5 <- splitForTrainingAndTest(surveyValues5, surveyTargets5, ratio=0.15)
```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
set.seed(4)
model5 <- mlp(survey_mlp5$inputsTrain, 
             survey_mlp5$targetsTrain, 
             size = c(10, 5), 
             initFunc = "Randomize_Weights",
             initFuncParams = c(-0.3, 0.3),
             learnFuncParams=c(0.1, 0),
             learnFunc = "Rprop",
             maxit = 100, 
             inputsTest=survey_mlp5$inputsTest, 
             targetsTest=survey_mlp5$targetsTest)


```


```{r size = "footnotesize", warning = FALSE, message = FALSE}
set.seed(45)
pred5 <- stats::predict(model5, newdata = survey_mlp5$inputsTest)
probs5 <- exp(pred5) / (1 + exp(pred5))
pMSE5 <- 1 / (2 * n) * sum((probs5 - 1 / 2)^2)
pMSE5

```

```{r}
(pMSE+pMSE2+pMSE3+pMSE4+pMSE5)/5

```






